A Plasma Torus Around a Young Low-Mass Star

### 2025-04-07 paper/ms.tex
update (latex cleaning)

% ** keck/hires: emission beyond v/veq>1
The spectroscopy shows emission beyond the star's equatorial velocity
($v_{\rm eq}$=130\,\kms).  There are at least two distinct emission
components, separated by half a cycle in phase.  The inner component at
lower velocities has clearer sinusoidal behaviour in time and is
double-peaked, with peak semi-amplitudes of 2.07\,$v_{\rm eq}$ and
2.88\,$v_{\rm eq}$ (see \S~\ref{subsec:halpha}).  There is significant
non-periodic variability in the emissivity of this double-peaked
component: the flux excess begins with an amplitude 70\% that of the
continuum early in the observation sequence, and diminishes to 10\% by
its end.  The higher-velocity component 180$^\circ$ opposite in phase is
detected only from $\phi$=0.2-1.0.  From $\phi$=0.2-0.5, this outer
component appears connected to the star in velocity space.  While its
peak semi-amplitude of 3.88\,$v_{\rm eq}$ is achieved at both
$\phi$=0.25 and 0.75, its amplitude similarly decreases from a 60\%
excess over the continuum at the beginning of the observation sequence
to a 10\% excess by its end.  The apparent period for all three
spectroscopic emission components is consistent with the photometric
\periodhr\ hour period.

@@

% * keck/hires: behavior at v/veq<1
The H$\alpha$ line core is more complex.  At $|\Delta v / v_{\rm
eq}|<1$, most observed H$\alpha$ photons come from the star's
chromosphere; circumstellar material might also modulate the line
profile.  Figure~\ref{fig:spec}e suggests line core variability caused
by both bright and dark regions on the star's surface, superposed with
smaller-amplitude variability from the transiting circumstellar
material.  For instance, from $\phi$=0-0.3, the double-peaked emission
feature is visible when viewed both on and off-limb; this feature is
circumstellar in origin.  However, the large emission feature that
crosses the star from $\phi$=0.4-0.9 emits at an amplitude greater than
that observed from the circumstellar components, and it crosses the
stellar velocity surface at a speed that suggests it instead originates
from a chromospherically bright region on the star's surface.
Similarly, from $\phi$=0.6-1.15 a 20\% deep absorption feature slowly
crosses the H$\alpha$ line profile.  This feature suggests either a
chromospherically dark region (e.g.~a starspot group) crossing the
star's surface, or an azimuthally extended absorptive component to the
circumstellar material which is not visible in emission.  Although the
origin of the other bright and dark streaks passing across the line core
are similarly ambiguous, a final exercise to quantify the behavior of the
line core is shown in Figure~\ref{fig:spec}d, where $f_{\rm H\alpha\
core}$ denotes the summed flux at $|\Delta v / v_{\rm eq}|<1$.  This
panel shows that changes in the line core flux correlate with the
broadband variability throughout most of the light curve, except near
$\phi$$\approx$0.5, corresponding to the transit of the 3.9\,$R_\star$
clump and the occultation of the lower-velocity clump.

@@

\begin{figure}[!b]
  \centering
  \includegraphics[width=0.99\textwidth]{figures/sf1.pdf}
  \caption{Photometric evolution of TIC 141146667 near the Keck/HIRES
  observation (green bar). 
  {\bf a,} TESS simple aperture photometry.  The main data gaps were
  caused by scattered light from the Earth (BTJD 3356-3358.5) and Moon
  (BTJD 3359.5-3360.5).  Raw two minute data are in gray; black
  time-averages to ten minute sampling.
  {\bf b-c,} Folded TESS light curve before and after spectroscopy.  Raw
  two minute data are in gray; black phase-averages to 100 points per
  cycle.  During BTJD 3352-3356, a state switch occurred near BTJD
  3353, in which the dip at $\phi$$\approx$0.8 disappeared.  While the
  large eclipse feature was present both before and after the HIRES
  sequence, its photometric shape evolved during the data gap.
  The oranges bar denote times of spectroscopic transits for the inner
  two H$\alpha$ clumps observed with HIRES.
  }
  \label{fig:fulllc}
\end{figure}

@@

%\begin{table}
  %\small

@@

  %\hline 
  %\hline 
  %Time [BJD$_{\rm TDB}$] & RV (\kms) & $\sigma_{\rm RV}$ (\kms) \\
  % from
  % results/halpha_to_rv_timerseries/multigauss_parametertable.tex

@@

%\end{table}

@@

\item[Acknowledgments]
  LGB is grateful to M.~Jardine, A.~Weinberger, B.~Tofflemire, J.~Spake,
  and J.Winn for insightful conversations that significantly informed
  this work, and to A.~Howard and H.~Isaacson for their assistance
  reducing the HIRES spectra.
  This work was supported by the Carnegie Fellowship and by the
  Heising-Simons 51~Pegasi~b Fellowship.
  %
  Funding for the TESS mission is provided by NASA’s Science Mission
  directorate.
  TESS is a product of millions of hours of work by thousands of people;
  we thank the TESS team for their efforts to make the mission a
  continued success.
  The HIRES data were obtained at the Keck Observatory, which exists
  through a similar scale of community effort.
  We recognize the importance that the summit of Maunakea has always had
  within the indigenous Hawaiian community, and we are deeply grateful 
  for the opportunity to conduct observations from this mountain.

>>>>

The spectroscopy shows emission beyond the star's equatorial velocity
($v_{\rm eq}$=130\,\kms).  There are at least two distinct emission
components, separated by half a cycle in phase.  The inner component
at lower velocities has clearer sinusoidal behaviour in time and is
double-peaked, with peak semi-amplitudes of 2.07\,$v_{\rm eq}$ and
2.88\,$v_{\rm eq}$ (see \S~\ref{subsec:halpha}).  There is significant
non-periodic variability in the emissivity of this double-peaked
component: the flux excess begins with an amplitude 70\% that of the
continuum early in the observation sequence, and diminishes to 10\% by
its end.  The higher-velocity component 180$^\circ$ opposite in phase
is detected only from $\phi$=0.2-1.0.  From $\phi$=0.2-0.5, this outer
component appears connected to the star in velocity space.  While its
peak semi-amplitude of 3.88\,$v_{\rm eq}$ is achieved at both
$\phi$=0.25 and 0.75, its amplitude similarly decreases from a 60\%
excess over the continuum at the beginning of the observation sequence
to a 10\% excess by its end.  The apparent period for all three
spectroscopic emission components is consistent with the photometric
\periodhr\ hour period.

@@

The H$\alpha$ line core is more complex.  At $|\Delta v / v_{\rm
eq}|<1$, most observed H$\alpha$ photons come from the star's
chromosphere; circumstellar material might also modulate the line
profile.  Figure~\ref{fig:spec}e suggests line core variability caused
by both bright and dark regions on the star's surface, superposed with
smaller-amplitude variability from the transiting circumstellar
material.  For instance, from $\phi$=0-0.3, the double-peaked emission
feature is visible when viewed both on and off-limb; this feature is
circumstellar in origin.  However, the large emission feature that
crosses the star from $\phi$=0.4-0.9 emits at an amplitude greater
than that observed from the circumstellar components, and it crosses
the stellar velocity surface at a speed that suggests it instead
originates from a chromospherically bright region on the star's
surface.  Similarly, from $\phi$=0.6-1.15 a 20\% deep absorption
feature slowly crosses the H$\alpha$ line profile.  This feature
suggests either a chromospherically dark region (e.g.~a starspot
group) crossing the star's surface, or an azimuthally extended
absorptive component to the circumstellar material which is not
visible in emission.  Although the origin of the other bright and dark
streaks passing across the line core are similarly ambiguous, a final
exercise to quantify the behavior of the line core is shown in
Figure~\ref{fig:spec}d, where $f_{\rm H\alpha\ core}$ denotes the
summed flux at $|\Delta v / v_{\rm eq}|<1$.  This panel shows that
changes in the line core flux correlate with the broadband variability
throughout most of the light curve, except near $\phi$$\approx$0.5,
corresponding to the transit of the 3.9\,$R_\star$ clump and the
occultation of the lower-velocity clump.

@@

\begin{figure}[!b]
  \centering
  \includegraphics[width=0.99\textwidth]{figures/sf1.pdf}
  \caption{Photometric evolution of TIC 141146667 near the Keck/HIRES
  observation (green bar).  {\bf a,} TESS simple aperture photometry.
  The main data gaps were caused by scattered light from the Earth
  (BTJD 3356-3358.5) and Moon (BTJD 3359.5-3360.5).  Raw two minute
  data are in gray; black time-averages to ten minute sampling.  {\bf
  b-c,} Folded TESS light curve before and after spectroscopy.  Raw
  two minute data are in gray; black phase-averages to 100 points per
  cycle.  During BTJD 3352-3356, a state switch occurred near BTJD
  3353, in which the dip at $\phi$$\approx$0.8 disappeared.  While the
  large eclipse feature was present both before and after the HIRES
  sequence, its photometric shape evolved during the data gap.  The
  oranges bar denote times of spectroscopic transits for the inner two
  H$\alpha$ clumps observed with HIRES.}
  \label{fig:fulllc}
\end{figure}

@@

@@

  % from results/halpha_to_rv_timerseries/multigauss_parametertable.tex

@@

@@

\item[Acknowledgments]
  LGB is grateful to M.~Jardine, A.~Weinberger, B.~Tofflemire,
  J.~Spake, J.~Winn, and L.~Hillenbrand for insightful conversations
  that significantly informed this work, and to A.~Howard and
  H.~Isaacson for their assistance reducing the HIRES spectra.
  This work was supported by the Carnegie Fellowship and by the
  Heising-Simons 51~Pegasi~b Fellowship.
  %
  Funding for the TESS mission is provided by NASA’s Science Mission
  directorate.
  TESS is a product of millions of hours of work by thousands of people;
  we thank the TESS team for their efforts to make the mission a
  continued success.
  The HIRES data were obtained at the Keck Observatory, which exists
  through a similar scale of community effort.
  We recognize the importance that the summit of Maunakea has always had
  within the indigenous Hawaiian community, and we are deeply grateful 
  for the opportunity to conduct observations from this mountain.
